### 2023-05-04 Documentation/UFMG/PFCII/Introduction/Introduction.tex
documentation

\chapter{Introduction}
\label{chap:intro} 
%este label será usado para referenciar este capítulo

Electrohydrodynamic Atomization (EHDA) is a way to disintegrate a liquid into droplets by exposing it to a strong electric field.\cite{prunet}
The balance between capillary forces and the electric field on the charged liquid defines the spraying dynamics and droplet size.
The electric current transported by the spray reveals characteristic shapes for different spray modes.
Signal processing techniques can allow a non-visual classification of the spray mode based on the electric current shape.\cite{Sjaaks}
The spray process imposes noise and random sequences on the measured signal making its classification not a trivial task. 
Industrial applications demand automated stabilization of a spray mode. 
This can be achieved by a closed-loop control system. 
This project is about to develop an application that can classify what dynamics the EHDA experiment is current in and control the variables to stabilize in the desired mode.

\section{Motivation and Justify}
\label{sec:motivacao}

% Argumente sobre a importância do projeto desenvolvido usando uma visão de alto nível, sem entrar em detalhes. Contextualize seu projeto dentro do local de execução ou da literatura e explique como ele é necessário ou inovador. É possível fazer uma breve revisão bibliográfica, confrontando seu trabalho com outras referências bibliográficas para mostrar a sua contribuição. No quesito contribuição, é muito importante deixar claro o tempo todo que partes do projetos foram executadas por outros e que partes foram executadas por você. Caso contrário, corre-se o risco de inadvertidademente tomar crédito pelo trabalho de outrem, o que pode ter implicações legais.

EHDA research has contributed as an important tool for the development of technology. 
The advantage of using EHDA is precision and uniform size and shape of droplets creation. Specially in certain spraying modes.

@@

The flowing electrical current carried by the spray reveals characteristic shapes for different atomization modes.
Signal processing techniques may allow non-visual classification of spray mode based on electrical current form.

The spraying process imposes noise and random sequences on the measured signal making it sorting is not a trivial task.
Industrial applications require automated stabilization of a spray mode. 
This can be achieved by a closed system loop control system. 
Automated spray mode sorting is a crucial part of a control system, as well as the development of an appropriate control algorithm.

@@

This project aims to give continuity to the previous student work\cite{Monica}, 
Mônica, who focused in detecting undesired discharges (sparks) in the system accusing high electric potential. 
For that, she developed a python software routine to connect most of the peripherals and analyzed the current data.
Her work corroborates the validation of this project motivating its development and optimization.

Above is shown the main goals listed:

\begin{enumerate}[]
\item Multipurpose applications for both scientific and industrial approaches.
\item Fully automated and intuitive system for EHDA.
\item Real Time non-visual classification of spraying modes.
\item Control and stabilization on a desired dynamic. 
\item System portability and versatility.
\end{enumerate}

@@

The NHL Stenden Water Technology group has been involved in previous projects that have successfully implemented automated signal processing techniques, resulting in highly ranked outcomes. 
However, further research is required to enhance the accuracy of the classification algorithm. In order to achieve this, the Water Technology Group at NHL Stenden University of Applied Sciences, in collaboration with Dutch companies, is conducting extensive research and implement appropriate classification algorithms. The aim is to combine analytical capabilities with infrastructure knowledge and availability to achieve optimal results.
As a student from UFMG, I am now actively involved in this research project to improve the automation usability, classification accuracy and system stabilization with signal processing techniques.

@@

>>>>

\chapter{Introduction}
\label{chap:intro}

Electrohydrodynamic Atomization (EHDA) is a way to disintegrate a liquid into droplets by exposing it to a strong electric field.\cite{prunet}
The balance between forces on the charged liquid meniscus defines the electrospraying dynamics and droplet size.
The electric current transported by the spray reveals characteristic shapes for different spray modes.
Signal processing techniques can allow a non-visual classification of the spray mode based on the electric current shape.\cite{Sjaaks}
The spray process imposes noise and random sequences on the measured signal making its classification a non-trivial task. 
Industrial applications demand automated stabilization of a spray mode. 
This can be achieved by a closed-loop control system. 
This project targets the development of a system which allows real time acquisition of the electrospray current, treat and process its values, define in which specific mode the electrospray is running and control the necessary hardware to change and/or stabilize the spray at a desired mode.

\section{Motivation and Justify}
\label{sec:motivation}

EHDA research has contributed as an important tool for the development of technology. 
The advantage of using electrospray is precision and uniform size and shape of droplets creation. 
Specially in certain spraying modes.

@@

Many applications such as gas odorization, spray coating and pharmaceutical industries require automated stabilization of a spray mode. 
This can be achieved by a closed system loop control system. 
An automated spray mode classification is a crucial part of the control system to work, as well as the development of an appropriate control algorithm.

@@

This project aims to give continuity to the previous student work\cite{Monica}, 
Mônica, who focused in detecting undesired discharges (sparks) in the system accusing high electric potential. 
For that, she developed a python software routine to connect most of the peripherals and analyzed the current data.
Her work corroborates the validation of this project motivating its continuation on development and optimization.

Bellow are shown the main goals listed:

\begin{enumerate}[]
\item Multipurpose applications for both scientific and industrial approaches.
\item Fully automated and intuitive system for EHDA.
\item Real time non-visual classification of electrospraying modes.
\item Control and stabilization on a desired spray. 
\item System portability and versatility.
\end{enumerate}

@@

The NHL Stenden Water Technology group has been involved in previous projects that have successfully implemented automated signal processing techniques, resulting in highly ranked outcomes. 
However, further research is required to enhance the accuracy of the classification algorithm.

In order to achieve this, the Water Technology Group at NHL Stenden University of Applied Sciences, in collaboration with Dutch companies, is conducting extensive research and implement appropriate classification algorithms. The aim is to combine analytical capabilities with infrastructure knowledge and availability to achieve optimal results.

As a student from UFMG, I am now actively involved in this research project to improve the automation usability, classification accuracy and system stabilization with signal processing techniques.

@@

This document is divided in 6 chapters.

Chapter \ref{chap:intro} provides an introduction to EHDA and explains the inspiration behind the project.

Chapter \ref{chap:lit_review} delves into the literature concepts and knowledge that were utilized in this project.

Chapter \ref{chap:system_description} describes the system that was implemented to make this project work, along with the instruments and models used to apply our methodology.

Chapter \ref{chap:Methodology} presents the methodology applied in our experiments to validate our system.

In Chapter \ref{chap:Results}, we present our results and engage in discussions comparing the knowledge acquired through literature review.

Finally, in Chapter \ref{chap:conclusion}, we conclude our document with discussions about the goals achieved and areas for potential optimization.

### 2023-05-04 Documentation/UFMG/PFCII/LiteratureReview/LiteratureReview.tex
documentation

\begin{figure}[H]
  \centering
  \resizebox{90mm}{!}{\includegraphics{Figuras/system_setup.png}}
  \caption{EHDA experiment setup \cite{Luewton}}
  \label{fig:ehda_setup}
\end{figure}

@@

  The existing signal contains valuable data not only about dripping but also other modes of spraying, thus it can serve as a non-visual classifier.
  This report draws significant inspiration from Sjaak's\cite{Sjaaks} work, which identifies distinct statistical characteristics of electrical current for various spraying modes.
  A profound use of his method will be explained in section \ref{sec:section_classification}.
  Together with the current characteristics, visual observations and results from literature it was investigated whether generic trends are present that can be related to the actual spraying modes.

>>>>

\begin{figure}[H]
  \centering
  \resizebox{100mm}{!}{\includegraphics{Figuras/system_setup.png}}
  \caption{EHDA experiment setup \cite{Luewton}}
  \label{fig:ehda_setup}
\end{figure}

@@

  The existing signal contains valuable data not only about dripping but also other modes of spraying, thus it can serve as a non-visual classifier.
  This project draws significant inspiration from Sjaak's\cite{Sjaaks} work, which identifies distinct statistical characteristics of electrical current for various spraying modes.
  A profound use of his method will be explained in section \ref{sec:section_classification}.
  Together with the current characteristics, visual observations and results from literature it was investigated whether generic trends are present that can be related to the actual spraying modes.

### 2023-05-04 Documentation/UFMG/PFCII/Results/Results.tex
documentation

The automated experiment routine is capable of acquiring a significantly more precise and extensive amount of data than what can be achieved by a human. 
With the save data by streaming in saving thread\ref{subsec:save_thread}, we achieved less program memory during the experiment and, specially, safety for not losing the data in case the program fail during an experiment.

Also, most power supplies rely on manual potentiometer adjustment to select the set point, which results in imprecise potential selection by a person. 
Moreover, manual adjustment takes some time to achieve the desired voltage. 
By running a computer routine that automatically sends commands to the power supply and pump machine we can achieve faster and more reliable experiment data points.

The visual interface for the user, seen in figures \ref{fig:multi_class_exp1} and \ref{fig:multi_class_exp}, shows all the important sensor data and signal analysis that can be interesting to the operator in real time.

@@

\begin{figure}[H]
    \center
    {\includegraphics[width=16cm]{Figuras/19:03/axs2.png}}
    \caption{Print screen during an experiment. At this moment we are in Multi Jet spraying mode. At the left we can see the current signal has a smooth and constant curve. We can also see the automatic classification on the right side.}
    \label{fig:multi_class_exp}
\end{figure}

@@

@@

Our first classification results were made exploring just the voltage ranges. For that stabilized all the other parameters such as humidity, temperature and external noise.
First, we fixed a flow rate to 0.7 uL/min. and ran a step routine as defined in \ref{subsec:step_routine}. 
The figure \ref{fig:step_class} shows three graphs.

The first is the input voltage step graph. The range is between 3K-10K Volts. Step size of 50V and step time of 5 seconds.

The second is the output raw data collected in the voltage range. The sampling rate of the oscilloscope is 10KHz. Therefore, this graph has 1.5 Million data points.

The Third graph is the same the second but after the classification procedure.

% \begin{figure}[H]
%     \center
%     \includegraphics[width=12cm]{Figuras/19:03/voltage_step.png}
%     \caption{Input voltage step graph. The range is between 3K-10K Volts. Step size of 50V and step time of 5 seconds.}
%     \label{fig:step_voltage}
% \end{figure}

% \begin{figure}[H]
%     \center
%     \includegraphics[width=12cm]{Figuras/19:03/raw-data-example.png}
%     \caption{Output raw data collected in the voltage range experiment showed in \ref{fig:step_voltage}. The sampling rate of the oscilloscope is 10KHz. This graph has 1.5 Million data points.}
%     \label{fig:raw_data}
% \end{figure}

% \begin{figure}[H]
%     \center
%     \includegraphics[width=12cm]{Figuras/19:03/classified-data-example.png}
%     \caption{The same graph as \ref{fig:raw_data} but after the classification procedure. The colors are green: Dripping, Blue: Intermittent, Red: Cone Jet, Purple: Multi Jet and Black: Undefined.}
%     \label{fig:class_step_data}
% \end{figure}

\begin{figure}[H]
    \center
    \includegraphics[width=18cm]{Figuras/may/step_class.png}
    \caption{Automatic electrospray classification through the step routine.}
    \label{fig:step_class}
\end{figure}

@@

    \begin{figure}[H]
        \center
        \includegraphics[width=8cm]{Figuras/regions.png}
        \label{fig:stability_1}
        \caption{Spraying modes regions of pure ethanol}
    \end{figure}

@@

        \begin{figure}[H]
            \center
            \includegraphics[width=17cm]{Figuras/19:03/control_first_results.png}
            \caption{Simple controller validation experiment. For this experiment we waited the system to stabilize in the desired spraying mode. After that we made a perturbation robustness test changing the flow rate.}
        \end{figure}

>>>>

The automated experiment routine is capable of acquiring a significantly more precise and extensive amount of data than what can be achieved by a human. 
With the save data by streaming in saving thread \ref{subsec:save_thread}, we achieved less program memory during the experiment and, specially, safety for not losing the data in case the program fail during an experiment.

Also, most power supplies rely on manual potentiometer adjustment to select the set point, which results in imprecise potential selection by a person and takes some time to achieve the desired voltage.
Moreover, the electrospray phenomena has a known hysteresis, that can be seen in figure \ref{fig:ganan_calvo_fig}, and can perform different results depending on the previous electric potential.
By running a computer routine that automatically sends commands to the power supply and pump machine we can achieve faster and more reliable experiment data points.

The visual interface for the user, seen in figure \ref{fig:multi_class_exp1}, shows all the important sensor data and signal analysis that can be interesting to the operator in real time, serving as a supervisory system.

@@

@@

We will categorize our classification results into two main groups - the step routine and the map routine. These two routines have showed valuable insights and comprehensible results of the electrospraying process.

@@

Our first classification results were made exploring the voltage ranges. 
For that, we fixed a flow rate to 0.7 uL/min. and ran a step routine as defined in \ref{subsec:step_routine}. 
The figure \ref{fig:step_class} shows three graphs.

The first shows the controller output signal as an input voltage of the process. As it is a step routine we implemented a increasing voltage with steps of sizes 50V and time between each step of 5 seconds. The voltage range is between 3K~10K Volts.

The second is the raw output data collected by the oscilloscope in \emph{data\_acquisition\_thread()}. 
The sampling rate is 10KHz. Therefore, this experiment of 700s has 7 Million data points just of current data. 
This is an example of how scalable the data collected can be depending on the experiment time. This will be even more noticeable in mapping experiments.

The Third graph is the same data as the second after the classification procedure done by \emph{data\_processing\_thread()}.

\begin{figure}[H]
    \center
    \includegraphics[width=15cm]{Figuras/may/step_class.png}
    \caption{Automatic electrospray classification through the step routine.}
    \label{fig:step_class}
\end{figure}

@@

    \begin{figure}[H]
        \center
        \includegraphics[width=8cm]{Figuras/regions.png}
        \label{fig:stability_1}
        \caption{Experimental spraying modes regions of pure ethanol}
    \end{figure}

@@

        \begin{figure}[H]
            \center
            \includegraphics[width=17cm]{Figuras/19:03/control_first_results.png}
            \caption{Simple controller validation experiment. For this experiment we waited the system to stabilize in the desired spraying mode. After that we made a perturbation robustness test changing the flow rate.}
            \label{fig:control_results}
        \end{figure}

### 2023-05-04 Documentation/UFMG/PFCII/systemDescription/systemDescription.tex
documentation

\begin{figure}[H]
  \centering
  \resizebox{150mm}{!}{\includegraphics{Figuras/april/control_model.png}}
  \caption{EHDA automation system setup}
  \label{fig:control_model_fig}
\end{figure}

@@

    \subsection{Video Thread}

        Normally deactivated, that thread is responsible for triggering the camera in case we want to save a video of that sample.

    \subsection{Plot}

        The only running function that is not a thread because of the plotting library \emph{matplotlib} incompatibilities of running outside the main function. 
        It is responsible for plotting in real time the current sample acquired, and its respective fast Fourier transform to evaluate the sample frequency spectrum.
        It takes as input the \emph{plot\_data\_queue} from the \emph{processing\_thread()} and displays three graphs updated in real time on the screen during the experiment.

@@

    In this chapter we described the system description detailing the hardware instruments, how it was modelled and its implementation in the software. Figure \ref{fig:metodology_example1} shows a print screen of the user interface during an experiment.

    \begin{figure}[H]
      \centering
      \resizebox{150mm}{!}{\includegraphics{Figuras/report4/experiment_print1.png}}
      \caption{EHDA automation system setup}
      \label{fig:metodology_example1}
  \end{figure}
\clearpage

>>>>

\begin{figure}[H]
  \centering
  \resizebox{150mm}{!}{\includegraphics{Figuras/control_loop.png}}
  \caption{EHDA automation system setup}
  \label{fig:control_model_fig}
\end{figure}

@@

    \subsection{Plotting}

        The only running function that should run in the main thread because the plotting library \emph{matplotlib} incompatibilities of running outside the main function. 
        It is responsible for plotting in real time the current sample acquired, and its respective fast Fourier transform to evaluate the sample frequency spectrum.
        It takes as input the \emph{plot\_data\_queue} from the \emph{processing\_thread()} and displays three graphs updated in real time on the screen during the experiment.

@@

    In this chapter we described the system description detailing the hardware instruments, how it was modelled and its implementation in the software. 
    Figure \ref{fig:Methodology_example1} shows a print screen of the system being used.

    \begin{figure}[H]
      \centering
      \resizebox{150mm}{!}{\includegraphics{Figuras/report4/experiment_print1.png}}
      \caption{EHDA automation system setup}
      \label{fig:Methodology_example1}
  \end{figure}
\clearpage
